Paper 1

### 2022-06-13 scripts/Paper 1.Rmd
add months to rainfall

```{r Precipitation, echo=FALSE, message = FALSE, fig.width = 10, fig.height = 4}
input.daily %>% mutate(Precip = replace_na(Precip, 0)) %>%
  ddply(c("Year", "month"), summarise,
  monthly.total = sum(Precip)) %>%
  ggplot(., aes(x = month, y = monthly.total, group = Year, fill = Year)) +
       scale_x_discrete(limits = month.abb) +
  geom_col(position = "dodge") + 
  ylab("Monthly rainfall (mm)")+
  xlab("Month")+
  theme_classic()

#below is precip data from liaweenee - it only goes up to Jan 2022 right now
rainfall_L <- read.csv("R:/SET/PlantSci/Shared Work Spaces/Ecology/Peoples files/Marion Nyberg/PhD/Rainfall/rainfall.csv")
rainfall_L <- rainfall_L[-1]
colnames(rainfall_L)[4] <- "Precip"
rainfall_L$Year <- as.factor(rainfall_L$Year)
rainfall_L <- rainfall_L %>% filter(Year == 2019)
colnames(rainfall_L)[3] <- "day"
colnames(rainfall_L)[2] <- "month"
rainfall_L$Site <- "L"
rainfall_L$date <- as.Date(with(rainfall_L, paste(Year, month, day,sep="-")), "%Y-%m-%d")
rainfall_L$jday <- yday(rainfall_L$date)

@@

rainfall.SP <- input.daily %>% select(Precip, jday, month, Year)
rainfall.SP$Site <- "SP"
rainfall.SP$date <- as.Date(with(rainfall.SP, paste(Year, month, jday,sep="-")), "%Y-%m-%j")
rainfall.2merge <- rainfall_L %>% filter(Year == 2019  & jday >= 244 & jday < 354)

rainfall <- rbind.fill(rainfall.2merge, rainfall.SP)
rainfall$month <- as.factor(rainfall$month)

@@

rainfall.summary %>% 
  ggplot(., aes(x = month, y = monthly.sum, group = Year, fill = Site)) +
  geom_col(position = "dodge", aes(fill = Year)) + 
  ylab("Monthly rainfall (mm)")+
  xlab("month")+
  theme_classic()
```

@@

NEE.summary %>%
  ggplot(., aes(jday, seasonal.sum)) +
  geom_line(aes(colour = Year))

@@

```{r create new dataframe for timeseries analysis}
rainfall.2merge <- rainfall %>% filter(Year == 2019  & jday < 354)
input.daily$Site <- "SP"

tsData <- rbind.fill(rainfall.2merge, input.daily) %>% 
  select(Year, month, jday, Precip, NEE_LT.gC, ER_LT.gC, GPP_LT.gC) %>%
  mutate(
    NEE2weeks = lag(NEE_LT.gC, n = 14, default = NA)
  )

tsData$date <- as.Date(with(tsData, paste(Year, month, jday,sep="-")), "%Y-%m-%d")
lm1 <- lm(NEE2weeks~Precip, tsData)
summary(lm1)

lm2 <- lm(NEE_LT.gC~Precip, tsData)
summary(lm2)
```

@@

>>>>

```{r Precipitation, echo=FALSE, message = FALSE, fig.width = 10, fig.height = 4}
input.daily %>% mutate(Precip = replace_na(Precip, 0)) %>%
  ddply(c("Year", "month"), summarise,
  monthly.total = sum(Precip)) %>%
  ggplot(., aes(x = month, y = monthly.total, group = Year, fill = Year)) +
       #scale_x_discrete(limits = month.abb) +
  geom_col(position = "dodge") + 
  ylab("Monthly rainfall (mm)")+
  xlab("Month")+
  theme_classic()

#below is precip data from liaweenee - it only goes up to Jan 2022 right now
rainfall_L <- read.csv("R:/SET/PlantSci/Shared Work Spaces/Ecology/Peoples files/Marion Nyberg/PhD/Rainfall/rainfall.csv")
rainfall_L <- rainfall_L[-1]
colnames(rainfall_L)[4] <- "Precip"
rainfall_L$Year <- as.factor(rainfall_L$Year)
rainfall_L <- rainfall_L %>% filter(Year == 2019)
colnames(rainfall_L)[3] <- "jday"
colnames(rainfall_L)[2] <- "month"
rainfall_L$Site <- "L"

@@

rainfall.SP <- input.daily %>% select(Precip, jday, month, Year)
rainfall.SP$Site <- "SP"

rainfall <- rbind(rainfall_L, rainfall.SP)
rainfall$month <- as.factor(rainfall$month) 
rainfall$month <- month.abb[rainfall$month]

rainfall %>%  mutate(Precip = replace_na(Precip, 0)) %>%
  ddply(c("Year", "month"), summarise,
  monthly.total = sum(Precip)) %>%
  ggplot(., aes(x = month, y = monthly.total, group = Year, fill = Year)) +
       scale_x_discrete(limits = month.abb) +
  geom_col(position = "dodge") + 
  ylab("Monthly rainfall (mm)")+
  xlab("Month")+
  theme_classic()

@@

rainfall.summary %>% 
  ggplot(., aes(x = month, y = monthly.sum, group = Year, fill = Site)) +
  geom_col(position = "dodge", aes(fill = Year)) + 
  ylab("Monthly rainfall (mm)")+
  xlab("month")+
  theme_classic()

rainfall.summary %>% 
  ggplot(., aes(x = month, y = monthly.sum, group = Year, fill = Site)) +
  geom_point(aes(color = Year, size = 2)) + 
  ylab("Monthly rainfall (mm)")+
  xlab("month")+
  theme_classic()

@@

@@

```{r cumulative fluxes}

cumulative.co2fluxes <- input %>%
  drop_na(Timestamp) %>% # some NAs in the Timestamp for some reason. Need to look into
  ddply(c("Year", "jday", "month", "season"), summarise,
        Fco2 = gofC.f(mean(Fco2, na.rm = TRUE)),
        ER = gofC.f(mean(ER_SOLO, na.rm = TRUE)),
        GPP = gofC.f(mean(GPP_SOLO, na.rm = TRUE)))

input.co2fluxes$GPP <- ifelse(input.co2fluxes$GPP < 0 , NA, input.co2fluxes$GPP) # for some reason there were values < 0 for GPP

input.co2fluxes$ERcum_year <- as.numeric(unlist(tapply(input.co2fluxes$ER, input.co2fluxes$Year, cumsum)))
input.co2fluxes$NEEcum_year <- as.numeric(unlist(tapply(input.co2fluxes$Fco2, input.co2fluxes$Year, cumsum)))
input.co2fluxes$GPPcum_year <- as.numeric(unlist(tapply(input.co2fluxes$GPP, input.co2fluxes$Year, cumsum)))

input.co2fluxes$GPP<- ifelse(is.na(input.co2fluxes$GPP), "0", input.co2fluxes$GPP) # replace NA with 0 for cumsum

(NEE <- plot_ly(input.co2fluxes, x = ~jday, y = ~NEEcum_year, split = ~Year, name = ~Year, type = 'scatter', mode = 'lines'))

(ER <- plot_ly(input.co2fluxes, x = ~jday, y = ~ERcum_year, split = ~Year, name = ~Year, type = 'scatter', mode = 'lines'))

(GPP <- plot_ly(input.co2fluxes, x = ~jday, y = ~GPPcum_year, split = ~Year, name = ~Year, type = 'scatter', mode = 'lines'))

@@

TSData <- rainfall[, 1] # ts data
decomposedRes <- decompose(tsData, type="mult") # use type = "additive" for additive components
plot (decomposedRes) # see plot below
stlRes <- stl(tsData, s.window = "periodic")
```
